\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-03-16 honors_thesis/journal_paper/paper.tex

%% PSO Stuff
\DeclareMathOperator{\URand}{U}
\DeclareMathOperator*{\argmin}{arg\;min}
\DeclareMathOperator*{\argmax}{arg\;max}
\DeclareMathOperator*{\arginf}{arg\;inf}
\DeclareMathOperator*{\argsup}{arg\;sup}
\providecommand{\ppos}{\ensuremath{\Vec{x}}}
\providecommand{\pvel}{\ensuremath{\Vec{v}}}
\providecommand{\nbest}{\ensuremath{\Vec{b}_N}}
\providecommand{\pbest}{\ensuremath{\Vec{b}_P}}
\providecommand{\constriction}{\ensuremath{\chi}}
\providecommand{\coeff}{\ensuremath{\phi}}
\providecommand{\obs}{\ensuremath{\Vec{\xi}}}

%SpecExPSO Stuff
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

@@

All of these parallel techniques use additional processors to increase the
number of particles in the swarm, either by adding individual particles or by
adding entire new subswarms.  The number of iterations of PSO that the
algorithms can perform is thus inherently limited by the time it takes to
evaluate the objective function---additional processors add more particles, not
more iterations.  However, there are many functions for which performing more
iterations with fewer particles is a better use of resources than performing
fewer iterations with more particles.

@@

At each iteration of the algorithm, the position $\ppos_t$ and velocity
$\pvel_t$ of each particle are updated as follows:
\begin{align}
\label{eq:velupdate}
	\pvel_{t+1} &=
		\constriction \left[ \pvel_t +
			\coeff_1\URand()\otimes(\pbest - \ppos_t) +
			\coeff_2\URand()\otimes(\nbest - \ppos_t)
		\right] \\
\label{eq:posupdate}
	\ppos_{t+1} &= \ppos_t + \pvel_{t+1}
\end{align}
where \( \URand() \) is a vector of random numbers drawn from a uniform
distribution, the \( \otimes \) operator is an element-wise vector
multiplication, $\pbest$ (called personal best) is the best position the
current particle has seen, and $\nbest$ (called neighborhood best) is the best
position the neighbors of the current particle have seen~\cite{bratton-sis07}.
\( \coeff_1 \), \( \coeff_2 \), and \( \constriction \) are parameters with
prescribed values required to ensure convergence (2.05, 2.05, and .73,
respectively)~\cite{clerc-tec02}.

>>>>

%% PSO Stuff
\DeclareMathOperator*{\argmin}{arg\;min}
\DeclareMathOperator*{\argmax}{arg\;max}
\DeclareMathOperator*{\arginf}{arg\;inf}
\DeclareMathOperator*{\argsup}{arg\;sup}
\providecommand{\pers}{\ensuremath{P}}
\providecommand{\neigh}{\ensuremath{N}}
\providecommand{\nURand}{\ensuremath{U^\neigh}}
\providecommand{\pURand}{\ensuremath{U^\pers}}
\providecommand{\ppos}{\ensuremath{\Vec{x}}}
\providecommand{\pvel}{\ensuremath{\Vec{v}}}
\providecommand{\nbest}{\ensuremath{\Vec{b}^\neigh}}
\providecommand{\pbest}{\ensuremath{\Vec{b}^\pers}}
\providecommand{\constriction}{\ensuremath{\chi}}
\providecommand{\ncoeff}{\ensuremath{\phi^\neigh}}
\providecommand{\pcoeff}{\ensuremath{\phi^\pers}}
\providecommand{\obs}{\ensuremath{\Vec{\xi}}}
\providecommand{\ofunc}{\ensuremath{f}}
\providecommand{\swarm}{\ensuremath{swarm}}

%SpecExPSO Stuff
\providecommand{\indic}{\ensuremath{I}}
\providecommand{\specvel}{\ensuremath{V}}
\providecommand{\specpos}{\ensuremath{X}}
\providecommand{\ln}{\ensuremath{ln}}
\providecommand{\rn}{\ensuremath{rn}}
\providecommand{\caseset}{\ensuremath{\mathcal{C}}}
\providecommand{\casedef}{\ensuremath{(\pbest:\nbest)}}
\providecommand{\casexn}{\ensuremath{(x:\neigh*)}}
\providecommand{\casexx}{\ensuremath{(x:x)}}
\providecommand{\casexl}{\ensuremath{(x:\ln)}}
\providecommand{\casexr}{\ensuremath{(x:\rn)}}
\providecommand{\casepn}{\ensuremath{(\pers*:\neigh*)}}
\providecommand{\casepl}{\ensuremath{(\pers*:\ln)}}
\providecommand{\casepr}{\ensuremath{(\pers*:\rn)}}
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

@@

The purpose of this paper is to consider pso parallelization strategies for
clusters of thousands of processors and functions for which a single
evaluation will take on the order of hours.
Our purpose is to explore the question ``since pso is often run with
a swarm of size of 100, what would one do if 500 or a few thousand processors were available?''

Previous parallel techniques use additional processors to increase the
number of particles in the swarm, either by adding individual particles or by
adding entire new sub-swarms but in all cases increasing the number of particles.  The number of iterations of PSO that the
algorithms can perform is thus inherently limited by the time it takes to
evaluate the objective function---additional processors add more particles, not
more iterations.
for many functions there comes a point of diminishing returns with respect to adding particles.
!!!! TODO WARNING WARNING MADE UP NUMBERS GET REAL ONES!! !!!!!
For the well-know benchmark function Sphere and as shown in Figure XXXXa increasing the swarm size from 20 to 120 has a significant effect on the 
speed of convergence. In this case, the swam goes below $1*10^5$ in x iteration with 20 particles, but 
achieves this same value in  y iterations with a swarms size of 120, the additional 100 particles.
However consider the change going from 2000 particles to 2100 particles. Here the additional 100 particles
reduce the number of iterations to achieve $1*10^5$ from x iterations to y.
Other benchmark function behave similarly. ans shown in fig xxxxb for ?? and xxxxc ?? (in all case 20 dimensions, average of 50 runs).
Certainly these benchmark functions do not prove that this will be the case for all functions,
but it does provide some evidence that for at least some functions, adding additional particles does not seem to have much benefit
after some point, some time earlier fig xx some time later. This phenomena has been commented on by others
in the form of recommended swarms sizes for various types of functions cite standard swarm paper.
\begin{quote}
Skip this??

However, there are many functions for which performing more
iterations with fewer particles is a better use of resources than performing
fewer iterations with more particles.  
\end{quote}

@@

At each iteration of the algorithm, the position $\ppos_t$ and velocity
$\pvel_t$ of each particle $i$ are updated as follows:
\begin{align}
\label{eq:velupdate}
	\pvel_{i,t+1} &=
		\constriction \left[ \pvel_t +
			\pcoeff\pURand_{i,t}\otimes(\pbest_{i,t} - \ppos_{i,t}) +
			\ncoeff\nURand_{i,t}\otimes(\nbest_{i,t} - \ppos_{i,t})
		\right] \\
\label{eq:posupdate}
	\ppos_{i,t+1} &= \ppos_{i,t} + \pvel_{i,t+1}
\end{align}
where \( \nURand_{i,t} \) and \( \nURand_{i,t} \) are vectors of i.i.d. random
numbers drawn from a uniform distribution, the \( \otimes \) operator is an
element-wise vector multiplication, $\pbest$ (called personal best) is the best
position the current particle has seen, and $\nbest$ (called neighborhood best)
is the best position the neighbors of the current particle have
seen~\cite{bratton-sis07}.  \( \ncoeff \), \( \pcoeff \), and \( \constriction
\) are parameters with prescribed values required to ensure convergence (2.05,
2.05, and .73, respectively)~\cite{clerc-tec02}. Note also that we have made
the particle identifier $i$ and the time step $t$ or $t+1$ explicit in the
subscript and moved other qualifiers, such as the personal $P$ and personal
best $P*$ qualifiers to the superscript position.  Superscripts can be miss
read as exponents, but they are qualifiers throughout this paper.
